---
title: "Combinatorial Capacity of English Negation in Child Language"
bibliography: library.bib
csl: apa6.csl
document-params: "10pt, letterpaper"

author-information: > 
    \author{{\large \bf Morton Ann Gernsbacher (MAG@Macc.Wisc.Edu)} \\ Department of Psychology, 1202 W. Johnson Street \\ Madison, WI 53706 USA
    \AND {\large \bf Masoud Jasbi (jasbi@ucdavis.edu)} \\ Department of Linguistics, 469 Kerr Hall, One Shields Avenue \\ Davis, CA 95 USA}

abstract: >
    Negation is very important for langauge and thought. How does it develop in the language of children? There has been many guessses like rejection, non-existence, denail, etc. but it has been hard to assess because these concepts are vaguge. Here we assess the combinatorial capacity of early negation in children's productions, and use words negation combines with as a proxy for early concepts expressed by it. We show some important stuff.
    
keywords: >
    Add your choice of indexing terms or keywords; kindly use a semi-colon; between each term.
    
output: cogsci2016::cogsci_paper
#final-submission: \cogscifinalcopy
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=3, fig.height=3, fig.crop = F, 
                      fig.pos = "tb", fig.path='figs/',
                      echo=F, warning=F, cache=F, 
                      message=F, sanitize = T)
```

```{r, libraries}
library(png)
library(grid)
library(ggplot2)
library(xtable)
library('RColorBrewer')
library(gridExtra)
library(tidyverse)
```

# Introduction

Negation is an abstract concept, lexicalized in all previously studied human languages, and crucial to everyday communication. It can help a coffee shop divide its menu into "coffee" and "not coffee" sections, with the "not coffee" section bringing together diverse items that otherwise cannot be labeled. It can help us regulate others' actions in a sign like "no mask, no entry".  It can also communicate our deepest wants and dislikes. But how does this crucial abstract concept emerge in humans? Does language play a role in its emergence or does language simply adopt it for communication? 

There has been several influential hypotheses on the conceptual origin of negation. 

In this paper, we address the same issues with a slightly different approach. We start with the widely accepted assumption that negation is a higher order operator or function, operating on lower level concepts. The question we ask is: what type of concepts does linguistic negation operate on in early child language? Do we find negation starting in a limited conceptual domain and then expanding to others? Or do we find it operating across different conceptual domains as early as we can attest it?

@darwin1872expression thought that negation has roots in the expression of human emotions and desires. He hypothesized that the earliest manifestation of negation and affirmation in infants is when they refuse food from parents, by withdrawing their heads laterally, or when they accept the food, by inclining their heads forward. He suggested that head shaking and nodding as common gestures for negation and affirmation have developed from this early habit. Considering early functions of negative morphemes like *no*, many researchers proposed that children use them to "reject" or "refuse" [@bloom1970language; @pea1978; @choi1988semantic]. For example, they may say "no" when asked "do you want juice?", say "not want it", or say "don't like it". @pea1978 proposed that this function of negation is the first to emerge in children.

Motor control: prohibition (do not spill milk), inability (I cannot zip it) 

@bloom1970language suggested that the first function of negation in children's speech is to express non-existence. Relates to children's development of object permanence. 


Perceptual: non-existence (no juice, no more milk, no fish in the bathroom, I do not have underpants), failure, Locatives (no in there, daddy was not on the phone), non-events (the dog not barking)

A third possible domain and path to the acquisition of negation is language itself. Word learning places its own constraints on the conceptual space. One possibility is that negation develops, and is aided by the act of labeling and categorizing objects and actions for linguistic communication. This function would manifest istelf in labeling acts with nominal predicates such as "this is not a bunny", "not red", or "this isn't a reptile". 

There has been no proposal for negation originating in the child's understanding of her own or other's epistemic states. In fact, most development theory of mind accounts assume that this ability emerges later in children. However, many corpus instances of negation modify mental state verbs such as *know*, *think*, and *remember* (e.g. "I not know"). Therefore, we also report the prevalence and emergence of such cases.  

Caveat on production vs comprehension. 

|Domain|Functions|Linguistic Composition|Examples|
  |------|--------|--------------------|
  |Emotion|Rejection| With *like* or *want* | "I not like it", "not want it" |
  |Motor Control| Prohibition, Inability | with imperative subjectless *do* or modal *can* | "do not spill milk", "I cannot zip it"|
  |Perception| non-existence | with a nominal, more-N, or locatives, event descriptions | no juice, no more milk, not in there, the dog not barking | 
  |Language Learning| Labeling | modifying predicative nominals | that's not a crocodile |
|Theory of Mind| Epistemic | know, think, remember  | I not know |
|Unknown | Temporal | now, today, again | not today |
|Unknown | Causal | why | why not? |
|Unknown | Posession | 's or pronouns, or have | not mine |

# Experiments

## Data and preprocessing 

For developmental data of child language in English, we turned to the CHILDES database [@macwhinney2000childes], which provides child-parent conversational interactions. We focused on utterances produced by children with typical development within the age range of 12 - 72 months, then extracted cases with any of the three negation markers that are of interest in this study: *no*, *not* and *n't*. Utterances with only one lexical token (e.g. *no !*) were not considered as here we aim to address particularly the question of what negation markers could *combine* with.
Preprocessing led to a data set of 365,260 utterances with negation structures from a total of 811 children across 56 corpora.

Figure \@ref(fig:speaker_stats)

```{r speakerdata, include=FALSE}

stats <-read.csv('data/age_info.txt', header = T, sep = '\t')
stats$Role[which(stats$Role == "Father")] <- 'Parent'
stats$Role[which(stats$Role == "Mother")] <- 'Parent'
stats$Role[which(stats$Role == "Target_Child")] <- 'Child'

child_stats <- data.frame(table(subset(stats, Role %in% c('Child'))$Age))
names(child_stats) <- c('Age', 'Freq')
child_stats$Ratio <- child_stats$Freq / sum(child_stats$Freq)
child_stats$Role <- rep('Child', nrow(child_stats))

parent_stats <- data.frame(table(subset(stats, Role %in% c('Parent'))$Age))
names(parent_stats) <- c('Age', 'Freq')
parent_stats$Ratio <- parent_stats$Freq / sum(parent_stats$Freq)
parent_stats$Role <- rep('Parent', nrow(parent_stats))

stats_data <- rbind(child_stats, parent_stats)

```

```{r speaker_stats, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3.5, fig.height=3.5, fig.cap = "Distribution of the number of utterances with negative morphemes in child and parent speech." }

stats_data %>% 
  ggplot(aes(as.numeric(Age), round(Ratio, 2), group = Role, color = Role)) +
  geom_point(aes(color = Role, shape = Role)) +
  scale_shape_manual(values = c(16, 3)) +
  scale_color_manual(values = c("forestgreen", "black")) + 
  geom_line(aes(color = Role, linetype = Role)) + 
  scale_x_continuous(breaks=seq(0, 72, 6)) +
  theme_classic() + 
  theme(text = element_text(size=10, family="Times")) + 
  theme(legend.position="top") +
  xlab("child age (months)") + 
  ylab("ratio")

```

## Negation concepts

In this section, we describe in details our automatic extraction of syntactic structures that express 
different types of negation concepts. The current English data from CHILDES contains morphosyntactic information [@sagae2010morphosyntactic] such as part-of-speech (POS) information as well as grammatical or syntactic dependency relations. We take advantage of information as such when automatically identifying our constructions of interest. An utterance with negative morpheme(s) was only considered when the negative morphemes has either a POS of *neg* or *qn*, the latter of which was mainly for cases with *no* as a quantifier. Furthermore, the syntactic functions and relations of the negative morphemes should not be enumeration (*no no no*), communicators or discourse markers

After extracting all instances with negative morphemes, the developmental trajectory of each construction type as described in the previous section was analyzed. While the matter of interest here is child speech, we also compared patterns in child production to those in parent speech as references at the corresponding age of the child. Then we combined the development of all construction types for analysis as a whole


### Emotion

In order to investigate utterances that express emotions, particularly the concept of rejections, we focused on specific cases where the head verb of the phrase is either *like* or *want* (including different forms of the verbs such as *liked* or *wanna* ), and the head verb is immediately preceded by one of the three negative morphemes. In particular, to not confuse with cases such as rhetorical questions (*don't you like it*) or declaratives (*you don't wanna do that*), which are more common in parent speech, our analyses were restricted to utterances where the head verb *like* or *want* takes either a subject *I* or has no subject. The existence of a subject was determined via searching for a word in the utterance that has the *SUBJ* dependency relation with the head verb. Overall our data extraction resulted in a total of 12,329 utterances (Child: 8,223; Parent: 4,106).
&nbsp;
(1) *like*: *I no like sea* 
(2) *want*: *no want one now* / *don't wanna go* 
&nbsp;

To compare the patterns between child and parent speech, for each domain, we calculated the relative ratio of (i) each of the three negative morphemes overall; (ii) usage of negative morphemes with the two different head verbs; (iii) utterances expression emotion with the three negative morphemes at different ages of the child. In both child and parent speech, when expressing emotion with either of the two head verbs *like* and *want*, the most frequently used negative morpheme is *n't* combined with an auxiliary verb. Comparing the two different head verbs, overall the negative morphemes are co-occur with *want* more frequently. 

```{r emotiondata, include=FALSE}

emotion <-read.csv('data/neg_emotion.txt', header = T, sep = '\t')
emotion <- subset(emotion, Subj == 'I' | Subj == 'NONE')
emotion$Role[which(emotion$Role == "Father")] <- 'Parent'
emotion$Role[which(emotion$Role == "Mother")] <- 'Parent'
emotion$Role[which(emotion$Role == "Target_Child")] <- 'Child'

child_emotion <- subset(emotion, Role %in% c('Child'))
parent_emotion <- subset(emotion, Role %in% c('Parent'))

child_like <- subset(child_emotion, Head %in% c('like'))
child_want <- subset(child_emotion, Head %in% c('want'))
parent_like <- subset(parent_emotion, Head %in% c('like'))
parent_want <- subset(parent_emotion, Head %in% c('want'))

child_like <- data.frame(table(child_like$Age))
child_like$Role <- rep('Child', nrow(child_like))
child_like$Head <- rep('like', nrow(child_like))
child_like$Ratio <- child_like$Freq / sum(child_like$Freq)

child_want <- data.frame(table(child_want$Age))
child_want$Role <- rep('Child', nrow(child_want))
child_want$Head <- rep('want', nrow(child_want))
child_want$Ratio <- child_want$Freq / sum(child_want$Freq)

parent_like <- data.frame(table(parent_like$Age))
parent_like$Role <- rep('Parent', nrow(parent_like))
parent_like$Head <- rep('like', nrow(parent_like))
parent_like$Ratio <- parent_like$Freq / sum(parent_like$Freq)

parent_want <- data.frame(table(parent_want$Age))
parent_want$Role <- rep('Parent', nrow(parent_want))
parent_want$Head <- rep('want', nrow(parent_want))
parent_want$Ratio <- parent_want$Freq / sum(parent_want$Freq)

emotion_data <- rbind(child_like, child_want, parent_like, parent_want)
names(emotion_data) <- c('Age', 'Freq', 'Role', 'Head', 'Ratio')

```

On the other hand, when looking at the developmental trajectory, as presented in Figure \@ref{fig:emotion}, children's usage of negative morphemes is comparable regardless of the particular head verb. In general, children did not start using negative morphemes in this domain until the age of 7 months; their usage of these morphemes for the concept of rejection is the most frequent during the age range of 13 - 21 months, then gradualy decreases as they age.


```{r emotion, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3.5, fig.height=3.5, set.cap.width=T, num.cols.cap=1, fig.cap = "Emotion" }

emotion_data %>% 
  ggplot(aes(as.numeric(Age), round(Ratio, 2), group = Role, color = Role)) +
  geom_point(aes(color = Role, shape = Role)) +
  scale_shape_manual(values = c(16, 3)) +
  scale_color_manual(values = c("steelblue", "peru")) + 
  geom_line(aes(color = Role, linetype = Role)) + 
#  facet_grid(. ~ Role) +
  facet_grid(vars(Head)) +
  scale_x_continuous(breaks=seq(0, 72, 6)) +
  theme_classic() + 
  theme(text = element_text(size=10, family="Times")) + 
  theme(legend.position="top") +
  xlab("child age (months)") + 
  ylab("ratio")

```


### Motor control

For utterances that articulate the concept of motor control, we focused on two individual aspects with different communicative functions. The first one includes cases that indicate prohibition (*e.g.* (3)), and the second one contain cases that articulate inability (*e.g.* (4)). For the former, we analyzed cases where the negative morphemes are combined with the auxiliary verb *do* (*do*, *does*, *did*) and the auxiliary does not take any subject; whereas for the latter, we analyzed cases where the negative morphemes co-occur with the auxiliary *can* (*can* and *could*). For inability in particular, similarly to the extraction of utterances that express rejection, in order to not accidentally include utterances such as rhetorical questions or imperatives (*you can't do that*), we excluded instances where the subject is not *I*. This led to a subset of 64,801 utterances (Child: 20,129; Parent: 44,672).
&nbsp;
(3) *do*: *don't blame Charlotte* 
(4) *can*: *I can't see* 
&nbsp;

Again for comparison of child and parent speech, we calculated the relative ratio of (i) each of the three negative morphemes overall; (ii) usage of negative morphemes with the two different communicative functions; (iii) utterances expression motor control with the three negative morphemes at different ages of the child. Overall the most frequently used negative morpheme is *n't* when applied in the domain of motor control. Comparing the two communicative functions, the negative morphemes tend to co-occur more often when expressing inability.

```{r motordata, include=FALSE}

motor <-read.csv('data/neg_motor.txt', header = T, sep = '\t')
motor <- subset(motor, Subj == 'I' | Subj == 'NONE')
motor$Role[which(motor$Role == "Father")] <- 'Parent'
motor$Role[which(motor$Role == "Mother")] <- 'Parent'
motor$Role[which(motor$Role == "Target_Child")] <- 'Child'

child_motor <- subset(motor, Role %in% c('Child'))
parent_motor <- subset(motor, Role %in% c('Parent'))

child_prohibition <- subset(child_motor, Function %in% c('prohibition') & Subj == 'NONE')
child_inability <- subset(child_motor, Function %in% c('inability'))
parent_prohibition <- subset(parent_motor, Function %in% c('prohibition') & Subj == 'NONE')
parent_inability <- subset(parent_motor, Function %in% c('inability'))

child_prohibition <- data.frame(table(child_prohibition$Age))
child_prohibition$Role <- rep('Child', nrow(child_prohibition))
child_prohibition$Function <- rep('prohibition', nrow(child_prohibition))
child_prohibition$Ratio <- child_prohibition$Freq / sum(child_prohibition$Freq)

child_inability <- data.frame(table(child_inability$Age))
child_inability$Role <- rep('Child', nrow(child_inability))
child_inability$Function <- rep('inability', nrow(child_inability))
child_inability$Ratio <- child_inability$Freq / sum(child_inability$Freq)

parent_prohibition <- data.frame(table(parent_prohibition$Age))
parent_prohibition$Role <- rep('Parent', nrow(parent_prohibition))
parent_prohibition$Function <- rep('prohibition', nrow(parent_prohibition))
parent_prohibition$Ratio <- parent_prohibition$Freq / sum(parent_prohibition$Freq)

parent_inability <- data.frame(table(parent_inability$Age))
parent_inability$Role <- rep('Parent', nrow(parent_inability))
parent_inability$Function <- rep('inability', nrow(parent_inability))
parent_inability$Ratio <- parent_inability$Freq / sum(parent_inability$Freq)

motor_data <- rbind(child_prohibition, child_inability, parent_prohibition, parent_inability)
names(motor_data) <- c('Age', 'Freq', 'Role', 'Function', 'Ratio')

```


As shown in Figure \@ref{fig:motor_control}, the developmental trajectory of using the negative morphemes in the domain of motor control is similar to that in the domain of emotion. Children started combining negative morphemes in syntactic structures that epxress prohibiiton and inability around the age of 6 months, then gradually increases until the age of 22 months.


```{r motor_control, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3.5, fig.height=3.5, set.cap.width=T, num.cols.cap=1, fig.cap = "Motor control" }

motor_data %>% 
  ggplot(aes(as.numeric(Age), round(Ratio, 2), group = Role, color = Role)) +
  geom_point(aes(color = Role, shape = Role)) +
  scale_shape_manual(values = c(16, 3)) +
  scale_color_manual(values = c("steelblue", "peru")) + 
  geom_line(aes(color = Role, linetype = Role)) + 
  facet_grid(vars(Function)) +
  scale_x_continuous(breaks=seq(0, 72, 6)) +
  theme_classic() + 
  theme(text = element_text(size=10, family="Times")) + 
  theme(legend.position="top") +
  xlab("child age (months)") + 
  ylab("ratio")

```


### Language learning

Within the domain of language learning, we concentrated on cases where negative morphemes are adopted to for labeing in the identity (*e.g.* (5) & (6)) or existence (*e.g.* (7)) of the nominal object that is being described. To do this, we extracted utterances that fit the following criteria: first, the negative morpheme has to be modifying a copula verb; secondly, the copula verb has to have a predicative nominal, which was identified with the help of POS information and dependency relation. The POS of the predicate has to be noun (*n*) and its dependency relation with the copula has to be *PRED*. This resulted in a total of 12,739 utterances (Child: 3,350 utterances; Parent: 9,389 utterances).
&nbsp;
(5) *that's not a farmer* 
(6) *I'm not a heavy baby Mum* 
(7) *can*: *there are no trees* 
&nbsp;

Again for comparison of child and parent speech, we calculated the relative ratio of (i) each of the three negative morphemes overall; (ii) usage of negative morphemes with the two different communicative functions; (iii) utterances expression motor control with the three negative morphemes at different ages of the child. Overall the most frequently used negative morpheme is *n't* when applied in the domain of motor control. Comparing the two communicative functions, the negative morphemes tend to co-occur more often when expressing inability.

```{r learningdata, include=FALSE}

learning <-read.csv('data/neg_learning.txt', header = T, sep = '\t')
learning$Role[which(learning$Role == "Father")] <- 'Parent'
learning$Role[which(learning$Role == "Mother")] <- 'Parent'
learning$Role[which(learning$Role == "Target_Child")] <- 'Child'

child_learning <- subset(learning, Role %in% c('Child'))
parent_learning <- subset(learning, Role %in% c('Parent'))

child_learning <- data.frame(table(child_learning$Age))
child_learning$Role <- rep('Child', nrow(child_learning))
child_learning$Function <- rep('labeling', nrow(child_learning))
child_learning$Ratio <- child_learning$Freq / sum(child_learning$Freq)

parent_learning <- data.frame(table(parent_learning$Age))
parent_learning$Role <- rep('parent', nrow(parent_learning))
parent_learning$Function <- rep('labeling', nrow(parent_learning))
parent_learning$Ratio <- parent_learning$Freq / sum(parent_learning$Freq)

learning_data <- rbind(child_learning,  parent_learning)
names(learning_data) <- c('Age', 'Freq', 'Role', 'Function', 'Ratio')

```


Comparing the three negative morphemes, the most frequently used is *not* in both child and parent speech. Based on results from Figure \@ref{fig:learning}, the developmental trajectory of using the negative morphemes in the domain of language learning is comparable to previous domains. Children started using the negative morphemes for the function of labeling nominal objects around the age of 7 months and their usage as such became more regular around the age of 10 months. However, the frequency of applying the negative morphemes in the language learning domain began to decrease around the age of 18 months.


```{r learning, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=3.5, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "Motor control" }

learning_data %>% 
  ggplot(aes(as.numeric(Age), round(Ratio, 2), group = Role, color = Role)) +
  geom_point(aes(color = Role, shape = Role)) +
  scale_shape_manual(values = c(16, 3)) +
  scale_color_manual(values = c("steelblue", "peru")) + 
  geom_line(aes(color = Role, linetype = Role)) + 
  facet_grid(vars(Function)) +
  scale_x_continuous(breaks=seq(0, 72, 6)) +
  theme_classic() + 
  theme(text = element_text(size=10, family="Times")) + 
  theme(legend.position="top") +
  xlab("child age (months)") + 
  ylab("ratio")

```

# Study

First level headings should be in 12 point , initial caps, bold and
centered. Leave one line space above the heading and 1/4~line space
below the heading.

## Method

Second level headings should be 11 point , initial caps, bold, and
flush left. Leave one line space above the heading and 1/4~ line
space below the heading.

### Third-Level Headings

Third-level headings should be 10 point , initial caps, bold, and flush
left. Leave one line space above the heading, but no space after the
heading.

# Discussion

Use standard APA citation format. Citations within the text should
include the author's last name and year. If the authors' names are
included in the sentence, place only the year in parentheses, as in
[-@NewellSimon1972a], but otherwise place the entire reference in
parentheses with the authors and year separated by a comma
[@NewellSimon1972a]. List multiple references alphabetically and
separate them by semicolons [@ChalnickBillman1988a; @NewellSimon1972a]. 
Use the et. al. construction only after listing all the authors to a
publication in an earlier reference and for citations with four or
more authors.

For more information on citations in RMarkdown, see **[here](http://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html#citations).**

## Footnotes

Indicate footnotes with a number\footnote{Sample of the first
footnote.} in the text. Place the footnotes in 9 point type at the
bottom of the page on which they appear. Precede the footnote with a
horizontal rule.\footnote{Sample of the second footnote.} You can also use 
markdown formatting to include footnotes using this syntax [^1].

[^1]: Sample of a markdown footnote.

## Figures

All artwork must be very dark for purposes of reproduction and should
not be hand drawn. Number figures sequentially, placing the figure
number and caption, in 10 point, after the figure with one line space
above the caption and one line space below it. If necessary, leave extra white space at
the bottom of the page to avoid splitting the figure and figure
caption. You may float figures to the top or bottom of a column, or
set wide figures across both columns.

## Two-column images

You can read local images using png package for example and plot 
it like a regular plot using grid.raster from the grid package. 
With this method you have full control of the size of your image. **Note: Image must be in .png file format for the readPNG function to work.**

You might want to display a wide figure across both columns. To do this, you change the `fig.env` chunk option to `figure*`. To align the image in the center of the page, set `fig.align` option to `center`. To format the width of your caption text, you set the `num.cols.cap` option to `2`.

```{r 2-col-image, fig.env = "figure*", fig.pos = "h", fig.width=4, fig.height=2, fig.align = "center", set.cap.width=T, num.cols.cap=2, fig.cap = "This image spans both columns. And the caption text is limited to 0.8 of the width of the document."}
img <- png::readPNG("figs/walrus.png")
grid::grid.raster(img)
```

## One-column images

Single column is the default option, but if you want set it explicitly, set `fig.env` to `figure`. Notice that the `num.cols` option for the caption width is set to `1`.

```{r image, fig.env = "figure", fig.pos = "H", fig.align='center', fig.width=2, fig.height=2, set.cap.width=T, num.cols.cap=1, fig.cap = "One column image."}
img <- png::readPNG("figs/lab_logo_stanford.png")
grid::grid.raster(img)
```


## R Plots

You can use R chunks directly to plot graphs. And you can use latex floats in the
fig.pos chunk option to have more control over the location of your plot on the page. For more information on latex placement specifiers see **[here](https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions)**

```{r plot, fig.env="figure", fig.pos = "H", fig.align = "center", fig.width=2, fig.height=2, fig.cap = "R plot" }
x <- 0:100
y <- 2 * (x + rnorm(length(x), sd = 3) + 3)

ggplot2::ggplot(data = data.frame(x, y), 
                aes(x = x, y = y)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```


## Tables

Number tables consecutively; place the table number and title (in
10 point) above the table with one line space above the caption and
one line space below it, as in Table 1. You may float
tables to the top or bottom of a column, set wide tables across both
columns.

You can use the xtable function in the xtable package.

```{r xtable, results="asis"}
n <- 100
x <- rnorm(n)
y <- 2*x + rnorm(n)
out <- lm(y ~ x)

tab1 <- xtable::xtable(summary(out)$coef, digits=c(0, 2, 2, 1, 2), 
                       caption = "This table prints across one column.")

print(tab1, type="latex", comment = F, table.placement = "H")
```

# Acknowledgements

Place acknowledgments (including funding information) in a section at
the end of the paper.

# References 

```{r}
# References will be generated automatically by Pandoc and included here.
# The following code is some latex to format the bibliography. Do not remove it.
```

\setlength{\parindent}{-0.1in} 
\setlength{\leftskip}{0.125in}
\noindent
