---
fontsize: 11pt
title: ""
# bibliography: ../negation_production_references.bib
output: pdf_document
---

\begin{center} \textbf{Parents' and Children's Production of English Negation} \end{center}

<!-- This gets rid of page numbers -->
\pagenumbering{gobble}

```{r packages, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggthemes)
library(wordcountaddin)
```

**Background** -- Previous research has proposed the following stages for children's production of negation (Klima \& Bellugi 1966; Wode 1977): 1. one-word negatives with *no*, 2. anaphoric *no* along with other words in responses to a previous utterance (e.g. no, outside) 3. pre-sentential non-anaphoric *no* or *not* (e.g. "no I see truck" to mean "I can't/don't see the truck") 4. *no*/*not* between the subject and the verb (e.g. "Kathrine no like celery"), adding *can't* and *don't* to the set of negators as unanalayzed wholes (e.g. "I can't open it") because *can* and *do* are not produced separately in this stage, 5. adult-like negation. However, stages 3 and 4 have been criticized for not being properly supported by the available data (de Villiers \& de Villiers 1979; Park 1979; Drozd 1995). In addition, Cameron-Faulkner et al. (2007) analyzed productions of a child in their corpus and argued for a "*no*-*not*-*n't* cline": children first produce *no*, followed by *not*, and finally the clitic form *n't*. This study uses the largest available child language corpus to bring more data to bear on these hypotheses.

```{r dataImport, include=FALSE}
negation_relfreq <- readRDS("../processed_data/category_table") 
negative_proportions <- readRDS("../processed_data/negation_proportions")
exclusions <- read_csv("../processed_data/exclusions.csv")

negative_proportions$category <- fct_recode(negative_proportions$category, `n't`="auxiliary", adverbial = "adverb", quantificational = "quantifier")

negative_proportions$category <- fct_relevel(negative_proportions$category, "no", "not", "n't")
```

**Method** -- We used the [CHILDES](https://childes.talkbank.org/) corpora (MacWhinney 2000), available via [childes-db](childes-db.stanford.edu) (Sanchez et al. 2019). English-North America and English-UK collections were tokenized. After exclusions due to unintelligibility or missing information (N = `r format(exclusions$Unintelligible + exclusions$missing + exclusions$age_ex, big.mark=",")`), the collection contained `r format(sum(negation_relfreq$count), big.mark=",")` tokens from the speech of `r exclusions$n_children` children. Tokens were classified into positive and negative morphemes, with the following negative morpheme categories: *no*, *not*, auxiliary clitic *n't*, negative pronouns and quantifiers (*none*, *nothing*, *nobody*, *no-one*, *nowhere*), negative adverb of frequency *never*, and derivational negatives with *un-* (e.g. unhappy), *in-* (e.g. invisible), *dis-*(e.g. disappear), *de-* (e.g. defrost), *non-* (e.g. nonsense), and *-less* (e.g. careless).

**Results** -- Between 12-18 months, almost all negative forms produced by children were instances of *no*, with some contracted auxiliary negatives like *don't* and *can't* (Figure 1). The proportions of *not* and its contracted form *n't* increased rapidly and simultaneously between 18-30 months until around 36-42 months they reached close to parent levels. Children started producing *not* and *n't* around the same time, if not *n't* slightly earlier. They started producing the positive auxiliaries such as *do* and *can* around the same time as negative ones (Figure 2). Compatible with Drozd's (1995) similar corpus findings, only a very small percentage of children's negative utterances could be analyzed as pre-sentential non-anaphoric negation.

**Conclusions** -- Results supported early productions of *no* (stage 1) before *not* and *n't*, but no evidence that *not* is produced before *n't*. We did not find support for robust non-anaphoric pre-sentential negation (stage 3) or contracted forms like *don't* and *can't* as unanalyzed wholes (stage 4). The results suggest a great degree of variability in children's productive development (de Villiers \& de Villiers 1979). Our findings are compatible with simultaneous development of frequent negative forms with a production bottleneck that favors shorter utterances like *no* to appear earlier.

```{r negationProportionPlot, echo=FALSE, fig.env="figure", fig.height=2.5, fig.cap="Proportion of different categories of negation in parents' and children's speech between 12 to 72 months of age (1 to 6 years)."}
negative_proportions %>%
  filter(age>13) %>%
  ggplot(aes(age,percent)) +
  geom_point(aes(color=category, shape=category)) +
  geom_line(aes(color=category)) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  facet_grid(.~speaker) +
  theme_classic() +
  theme(text = element_text(size=10, family="Times")) + xlab("child age (months)")
```

```{r auxRelFreq, echo=FALSE, fig.env="figure", fig.height=2.3, fig.cap="Relative frequency (per thousand words) of positive auxiliary (e.g. *do*, *can*, *is*, ...) as well as their contracted negatives (*don't*, *can't*, *isn't*, ...) in the speech of parents and children between 12 to 72 months of age (1 to 6 years)."}
negation_relfreq %>%
  filter(category=="auxiliary") %>%
  ggplot(aes(age,ppt)) +
  geom_point(aes(color=polarity, shape=polarity)) +
  scale_shape_manual(values=c(16, 3)) +
  geom_line(aes(color=polarity, linetype=polarity)) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  facet_grid(.~speaker) +
  theme_classic() + 
  theme(text = element_text(size=10, family="Times")) + xlab("child age (months)") + ylab("per 1000 words")
```

\footnotesize
\hfill Word Count: 497

**References** -- Cameron-Faulkner, Thea, Elena Lieven, and Anna Theakston (2007) What Part of "No" Do Children Not Understand? A Usage-Based Account of Multiword Negation. *Journal of Child Language (JCL)* -- de Villiers, Peter A, and Jill de Villiers (1979) Form and Function in the Development of Sentence Negation. *Papers and Reports on Child Language Development* -- Drozd, Kenneth (1995) Child English pre-sentential negation as metalinguistic exclamatory sentence negation. *JCL* -- Klima, Edward, and Ursula Bellugi (1966) Syntactic Regularities in the Speech of Children. *Psycholinguistics Papers* -- MacWhinney, Brian (2000) The CHILDES Project: The Database -- Park, Tschang-Zin (1979) Some Facts on Negation: Wode's Four-Stage Developmental Theory of Negation Revisited. *JCL* -- Sanchez, Alessandro, Stephan Meylan, Mika Braginsky, Kyle MacDonald, Daniel Yurovsky, and Michael C Frank (2019) Childes-Db: A Flexible and Reproducible Interface to CHILDES. *Behavior Research Methods* -- Wode, Henning (1977) Four Early Stages in the Development of L1 Negation. *JCL*
