---
title: "Negation Production Plots and Analyses"
author: "Masoud Jasbi"
date: "2/21/2018"
output: html_document
---

```{r load_library}
library(tidyverse)
library(magrittr)
library(feather)
library(binom)
library(childesr)
library(lubridate)
library(ggrepel)
```

```{r read_processed_data}
english_tokens <- read_feather("../processed_data/english_tokens_processed.feather")
```

```{r polarityTable}
polairty_table <-
  english_tokens %>%
  group_by(speaker, polarity, age) %>%
  summarize(count = n()) %>%
  group_by(speaker, age) %>%
  mutate(total = sum(count), proportion = count / total, percent = proportion * 100)
```

```{r polarityPlot}
polairty_table %>%
  filter(polarity=="negative") %>%
ggplot(aes(age, percent, color=speaker)) +
  geom_point(aes(color=speaker)) +
  geom_smooth(aes(fill = speaker), span=0.7) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic() + ggtitle("Production of Negative Words")
```

```{r categoryTable}
category_table <-
  english_tokens %>%
  group_by(speaker, category, polarity, age) %>%
  summarize(count = n()) %>%
  group_by(speaker, age) %>%
  mutate(total = sum(count), rel_freq = count / total, ppt = rel_freq * 1000)

saveRDS(negation_children, file="../processed_data/category_table")
```

```{r noNotPlot}
category_table %>%
  filter(category=="no" | category=="not") %>%
  ggplot(aes(age,ppt)) +
  geom_point(aes(color=speaker)) +
  facet_wrap(.~category) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  geom_smooth(aes(color=speaker), span=0.6) +
  theme_classic()
```

Positive and negative modal auxiliary forms:

```{r aux}
category_table %>%
  filter(category=="auxiliary") %>%
  ggplot(aes(age,ppt)) +
  geom_point(aes(color=polarity, shape=speaker)) +
  geom_line(aes(color=polarity, linetype=speaker)) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

```{r adverb}
category_table %>%
  filter(category=="adverb") %>%
  ggplot(aes(age,ppt)) +
  geom_point(aes(color=speaker)) +
  facet_wrap(.~polarity, scales="free_y") +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  geom_smooth(aes(color=speaker), span=0.7) +
  theme_classic()
```

```{r categoriesPlot}
category_table %>%
  filter(category!="other") %>%
  ggplot(aes(age,ppt)) +
  geom_point(aes(color=speaker)) +
  facet_wrap(polarity~category, scales="free_y") +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  geom_smooth(aes(color=speaker), span=0.7) +
  theme_classic()
```

```{r UnderTwoNegWords}
undertwo_negwords <-
  english_tokens %>%
  filter(age<25, speaker=="child", polarity=="negative", word!="other") %>%
  group_by(word) %>%
  summarize(count = n()) %>%
  filter(count>30) %>%
  arrange(desc(count))

undertwo_negwords
```

Let us assume an early representation of negation. The fact that "cant" appears suggests children can apply negation to representations of abilities. The fact that "didn't", "doesn't", "wont", "never", "isn't" appears suggests that negation can also operate over representations of time. "nothing" as well as many "no noun" instances show the role of negation in representing nonexistence. Many instances of "don't" also suggest negation can apply to representations of actions (don't run).

```{r negationProportion}
negative_proportions <-
  english_tokens %>%
  filter(polarity=="negative") %>%
  group_by(category, age, speaker) %>%
  summarize(count = n()) %>%
  group_by(age, speaker) %>%
  mutate(total = sum(count), proportion = count / total, percent = proportion * 100)

saveRDS(negative_proportions, file="../processed_data/negation_proportions")
```

```{r}
  english_tokens %>%
  filter(gloss=="all") %>%
  arrange(target_child_age)
```

# Proportion Plot

```{r negationPlot}
negative_proportions %>%
  ggplot(aes(age,percent)) +
  geom_point(aes(color=category, shape=category)) +
  geom_line(aes(color=category)) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  facet_grid(.~speaker) +
  theme_classic()
```

# Relative Frequencies

## Negation Categories

Relative frequency plot of major categories (no, not, n't) for parents and children:

```{r relativeFrequencyMajor}
category_table %>%
  filter(category!="other", category!="quantifier", category!="adverb", category!="derivational", polarity=="negative") %>%
ggplot(aes(age,ppt, fill=category, shape = category, color=category)) +
  geom_point(stat="identity", aes(fill=category, shape = category, color=category)) +
  geom_line(aes(fill = category)) +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

Relative frequency plot for minor categories (quantifiers, derivational, adverbs):

```{r relativeFreuencyMinor}
category_table %>%
  filter(category!="other", category!="no", category!="not", category!="auxiliary", polarity=="negative") %>%
ggplot(aes(age,ppt, fill=category, color=category, shape=category)) +
  geom_point(stat="identity", aes(fill=category, color=category, shape=category)) +
  geom_line(aes(fill = category)) +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

## Word Level Plots

Word level table:

```{r}
word_table <-
  english_tokens %>%
  group_by(speaker, word, category, polarity, age) %>%
  summarize(count = n()) %>%
  group_by(speaker, age) %>%
  mutate(total = sum(count), rel_freq = count / total, ppt = rel_freq * 1000)

saveRDS(word_table, file="../processed_data/word_table")
```

### Adverbials

Plot for adverbials below. It seems like "often" and "ususally" are not produced that often!

```{r AuxWordPlotDo}
word_table %>%
  filter(category=="adverb") %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line(aes(fill = word)) +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

Plot for "always" and "never" as positive and negative pairs:

```{r AuxWordPlotDo}
word_table %>%
  filter(word=="never" | word=="always") %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line(aes(fill = word)) +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

### Derivationals

The corpus is not big enough to give us good estimates of the derivational ones. *un-* and *dis-* seem to be the most common ones.

```{r AuxWordPlotDo}
word_table %>%
  filter(category=="derivational", age>13) %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line(aes(fill = word)) +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

### Quantifiers

Most common quantifier words are probably "some" and "something".

```{r AuxWordPlotDo}
word_table %>%
  filter(category=="quantifier") %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line(aes(fill = word)) +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

Plotting "none", "some", and "all"

```{r AuxWordPlotDo}
word_table %>%
  filter(word=="some" | word=="none") %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line(aes(fill = word)) +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```


```{r AuxWordPlotDo}
word_table %>%
  filter(word=="something" | word=="nothing" | word=="everything") %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line(aes(fill = word)) +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

some/no/every+body ones are too infrequent for this corpus to give us good estimates:

```{r AuxWordPlotDo}
word_table %>%
  filter(word=="everybody" | word=="nobody" | word=="somebody") %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line(aes(fill = word)) +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

some/no/every+hwere ones are the same:

```{r AuxWordPlotDo}
word_table %>%
  filter(word=="everywhere" | word=="nowhere" | word=="somewhere") %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line(aes(fill = word)) +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

### Auxiliaries

```{r AuxWordCant}
word_table %>%
  filter(word=="could" | word=="couldnt" | word=="can" | word=="cant" | word=="do" | word=="dont") %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line(aes(fill = word)) +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

```{r}
word_table %>%
  filter(word=="are" | word=="arent") %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line(aes(fill = word)) +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

```{r}
word_table %>%
  filter(word=="did" | word=="didnt") %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word), alpha=5/10) +
  geom_line(aes(fill = word, linetype=speaker)) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

# Percentage of Children Producing Negative Morphemes

Here I tried to test if we can find out about the proportion of children producing negative forms in CHILDES. However, it seems like this is not good practice. The database seems to be too sparse per child so we end up severly underestimating if a child is actually producing it or not. In order to infer what proportion of children produce a given form from naturalistic productions, we need a database that has at least comparable contribution from each child. This is not at all the case for CHILDES. 

```{r proportion_children}
proportion_children <-
  english_tokens %>%
  filter(speaker=="child") %>% #remove data from parents
  group_by(age) %>% #group by age and compute the total number of children in each age bin
  mutate(total_children=length(unique(target_child_id))) %>%
  group_by(negation_type, age, target_child_id, total_children) %>% #for each child, count their negation in each bin
  summarize(count = n()) %>%
  filter(count>2) %>% #remove children that had fewer than 5 tokens in the age bin
  group_by(negation_type,age) %>%
  mutate(producing_children=length(unique(target_child_id)), pc = producing_children/total_children)

proportion_children %>%
  filter(negation_type!="pos")%>%
  ggplot(aes(age, pc)) +
  geom_bar(stat="identity") +
  facet_wrap(.~negation_type)
```
