---
title: "Negation Production Plots and Analyses"
author: "Masoud Jasbi"
date: "2/21/2018"
output: html_document
---

```{r load_library}
library(tidyverse)
library(magrittr)
library(feather)
library(binom)
library(childesr)
library(lubridate)
library(ggrepel)
```

# Token Plots

```{r read_processed_data}
english_tokens <- read_feather("../processed_data/english_tokens_processed.feather")
```

## Proportion Plot

```{r negationProportion}
negative_proportions <-
  english_tokens %>%
  filter(polarity=="negative") %>%
  group_by(category, age, speaker) %>%
  summarize(count = n()) %>%
  group_by(age, speaker) %>%
  mutate(total = sum(count), proportion = count / total, percent = proportion * 100)

saveRDS(negative_proportions, file="../processed_data/negation_proportions")
```

```{r negationPlot}
negative_proportions %>%
  filter(!(age==12 & speaker=="child")) %>%
  filter(category !="quantifier", category !="derivational", category !="adverb") %>%
  ggplot(aes(age,percent, color=category)) +
  geom_point(aes(color=category, shape=category)) +
#  geom_line(aes(color=category)) +
  geom_smooth(span=1) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  facet_grid(.~speaker) +
  theme_classic()
```

```{r categoryTable}
category_table <-
  english_tokens %>%
  group_by(speaker, category, polarity, age) %>%
  summarize(count = n()) %>%
  group_by(speaker, age) %>%
  mutate(total = sum(count), rel_freq = count / total, ppt = rel_freq * 1000)

saveRDS(negation_children, file="../processed_data/category_table")
```

Positive and negative modal auxiliary forms:

```{r aux}
category_table %>%
  filter(category=="auxiliary") %>%
  ggplot(aes(age,ppt, color=polarity)) +
  geom_point(aes(color=polarity, shape=speaker)) +
#  geom_line(aes(color=polarity, linetype=speaker)) +
  facet_grid(.~speaker) +
  geom_smooth() +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

```{r categoriesPlot}
category_table %>%
  filter(category!="other") %>%
  ggplot(aes(age,ppt)) +
  geom_point(aes(color=speaker)) +
  facet_wrap(polarity~category, scales="free_y") +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  geom_smooth(aes(color=speaker), span=0.7) +
  theme_classic()
```

# Relative Frequencies

## Negation Categories

Relative frequency plot of major categories (no, not, n't) for parents and children:

```{r relativeFrequencyMajor}
category_table %>%
  filter(category!="other", category!="quantifier", category!="adverb", category!="derivational", polarity=="negative") %>%
ggplot(aes(age,ppt, fill=category, shape = category, color=category)) +
  geom_point(stat="identity", aes(fill=category, shape = category, color=category)) +
  geom_line(aes(fill = category)) +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

Relative frequency plot for minor categories (quantifiers, derivational, adverbs):

```{r relativeFreuencyMinor}
category_table %>%
  filter(category!="other", category!="no", category!="not", category!="auxiliary", polarity=="negative") %>%
ggplot(aes(age,ppt, fill=category, color=category, shape=category)) +
  geom_point(stat="identity", aes(fill=category, color=category, shape=category)) +
  geom_line(aes(fill = category)) +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

## Word Level Plots

Word level table:

```{r}
word_table <-
  english_tokens %>%
  group_by(speaker, word, category, polarity, age) %>%
  summarize(count = n()) %>%
  group_by(speaker, age) %>%
  mutate(total = sum(count), rel_freq = count / total, ppt = rel_freq * 1000)

saveRDS(word_table, file="../processed_data/word_table")
```

### Auxiliaries

```{r cantDont}
cando_plot <-
  word_table %>%
  filter(speaker== "child", word == "can" | word == "cant" | word == "do" | word == "dont")

cando_plot %>%
  filter(word == "do" | word == "dont") %>%
  ggplot(aes(age,ppt, color=word, shape=word)) +
  geom_point(aes(color=word, shape=word)) +
  geom_line(aes(color=word, shape=word)) +
#  geom_smooth() +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()

cando_plot %>%
  filter(word == "can" | word == "cant") %>%
  ggplot(aes(age,ppt, color=polarity, shape=word)) +
  geom_point(aes(color=polarity, shape=word)) +
  geom_line(aes(color=polarity, shape=word)) +
#  geom_smooth() +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

```{r AuxWordPlotDo}
aux_plot <-
  word_table %>%
  filter(category=="auxiliary", polarity=="negative", word!="aint", word!="other", word!="amnt", word!="maynt", word!="mightnt", word!="shant")
  
  
aux_plot$word <- fct_relevel(aux_plot$word, "dont", "cant", "didnt", "isnt", "doesnt", "wont", "arent", "wasnt", "havent", "hasnt", "couldnt", "wouldnt", "werent", "shouldnt", "hadnt", "mustnt")

  ggplot(aux_plot, aes(age, ppt, fill=speaker, color=speaker, shape=speaker)) +
  geom_point(stat="identity", aes(fill=speaker, color=speaker, shape=speaker)) +
  geom_line() +
  facet_wrap(.~word, scales = "free_y") +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  geom_vline(xintercept = 24, linetype = "dashed") +
  theme_classic()
```

Auxiliary words that are almost never said: maynt, mightnt, mustnt, 

```{r AuxWordPlotDo}
word_table %>%
  filter(word=="hadnt" | word=="hasnt" | word=="havent") %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line() +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

```{r AuxWordPlotDo}
word_table %>%
  filter(word=="didnt" | word=="doesnt" | word=="dont") %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line() +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

```{r AuxWordPlotDo}
word_table %>%
  filter(word=="cant" | word=="couldnt" | word=="shouldnt") %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line() +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

```{r AuxWordPlotDo}
word_table %>%
  filter(word=="isnt" | word=="arent") %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line() +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

```{r AuxWordPlotDo}
word_table %>%
  filter(word=="werent" | word=="wasnt") %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line() +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```


```{r AuxWordPlotDo}
word_table %>%
  filter(word=="wont" | word=="wouldnt") %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line() +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

### Adverbials

Plot for adverbials below. It seems like "often" and "ususally" are not produced that often!

```{r AuxWordPlotDo}
word_table %>%
  filter(category=="adverb") %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line(aes(fill = word)) +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

Plot for "always" and "never" as positive and negative pairs:

```{r AuxWordPlotDo}
word_table %>%
  filter(word=="never" | word=="sometimes" | word=="always") %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line() +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

### Derivationals

The corpus is not big enough to give us good estimates of the derivational ones. *un-* and *dis-* seem to be the most common ones.

```{r AuxWordPlotDo}
word_table %>%
  filter(category=="derivational", age>13) %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line(aes(fill = word)) +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

### Quantifiers

Most common quantifier words are "some" and "something".

```{r AuxWordPlotDo}
word_table %>%
  filter(category=="quantifier") %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line(aes(fill = word)) +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

Plotting "none", "some", and "all"

```{r AuxWordPlotDo}
word_table %>%
  filter(word=="some" | word=="none") %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line(aes(fill = word)) +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```


```{r AuxWordPlotDo}
word_table %>%
  filter(word=="something" | word=="nothing" | word=="everything") %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line(aes(fill = word)) +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

some/no/every+body ones are too infrequent for this corpus to give us good estimates:

```{r AuxWordPlotDo}
word_table %>%
  filter(word=="everybody" | word=="nobody" | word=="somebody") %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line(aes(fill = word)) +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

some/no/every+hwere ones are the same:

```{r AuxWordPlotDo}
word_table %>%
  filter(word=="everywhere" | word=="nowhere" | word=="somewhere") %>%
  ggplot(aes(age, ppt, fill=word, color=word, shape=word)) +
  geom_point(stat="identity", aes(fill=word, color=word, shape=word)) +
  geom_line(aes(fill = word)) +
  facet_grid(.~speaker) +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic()
```

## Utterance Plots and Analyses

```{r importProcessedUtterances}
eng_utterances <- read_csv("../processed_data/english_utterances_processed.feather")
```

```{r negativePercent}
negative_percent <-
  eng_utterances %>%
  group_by(speaker, age) %>%
  summarize(total= sum(no)+sum(not)+sum(nt), no = sum(no)/total, not=sum(not)/total, nt=sum(nt)/total) %>%
  gather("category", "percent", no:nt)
```

```{r NegativePercentPlot}
negative_percent %>%
  filter(percent!=0) %>%
  ggplot(aes(age, percent)) +
  geom_point(aes(color=category, shape=category)) + 
  geom_line(aes(color=category)) +
  facet_grid(.~speaker) +
  theme_classic()
```

```{r RelFreqs}
all_utterance_relfreqs <-
  eng_utterances %>%
  group_by(speaker, age) %>%
  summarize(no= sum(no)/sum(num_tokens)*1000, not=sum(not)/sum(num_tokens)*1000, nt=sum(nt)/sum(num_tokens)*1000) %>%
  gather("category", "ppt", no:nt)

saveRDS(all_utterance_relfreqs, "../processed_data/all_utterance_relfreqs")
```

```{r RelFreqPlot}
all_utterance_relfreqs %>%
  filter(ppt!=0) %>%
  ggplot(aes(age, ppt)) +
  geom_point(aes(color=category, shape=category)) + 
  geom_line(aes(color=category)) +
  facet_grid(.~speaker) +
  theme_classic()
```

```{r}
negative_percent_multi <-
  eng_utterances %>%
  filter(num_tokens>1, part_of_speech!="co co co", part_of_speech!="co co", part_of_speech!="co") %>%
  group_by(speaker, age) %>%
  summarize(total= sum(no)+sum(not)+sum(nt), no = sum(no)/total, not=sum(not)/total, nt=sum(nt)/total) %>%
  gather("category", "percent", no:nt)
```

```{r}
negative_percent_multi %>%
  filter(percent!=0) %>%
  ggplot(aes(age, percent)) +
  geom_point(aes(color=category, shape=category)) + 
  geom_line(aes(color=category)) +
  facet_grid(.~speaker) +
  theme_classic()

negative_percent_multi %>%
  filter(percent!=0) %>%
  ggplot(aes(age,percent, color=category)) +
  geom_point(aes(color=category, shape=category)) +
#  geom_line(aes(color=category)) +
  geom_smooth() +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  facet_grid(.~speaker) +
  theme_classic()
```

```{r}
multiword_relfreqs <-
  eng_utterances %>%
  filter(num_tokens>1, part_of_speech!="co co co", part_of_speech!="co co", part_of_speech!="co") %>%
  group_by(speaker, age) %>%
  summarize(no= sum(no)/sum(num_tokens)*1000, not=sum(not)/sum(num_tokens)*1000, nt=sum(nt)/sum(num_tokens)*1000) %>%
  gather("category", "ppt", no:nt)

saveRDS(multiword_relfreqs, "../processed_data/multiword_relfreqs")
```

```{r}
multiword_relfreqs %>%
  filter(ppt!=0) %>%
  ggplot(aes(age, ppt)) +
  geom_point(aes(color=category, shape=category)) + 
  geom_line(aes(color=category)) +
  facet_grid(.~speaker) +
  theme_classic()
```


# Percentage of Children Producing Negative Morphemes

Here I tried to test if we can find out about the proportion of children producing negative forms in CHILDES. However, it seems like this is not good practice. The database seems to be too sparse per child so we end up severly underestimating if a child is actually producing it or not. In order to infer what proportion of children produce a given form from naturalistic productions, we need a database that has at least comparable contribution from each child. This is not at all the case for CHILDES. 

```{r proportion_children}
proportion_children <-
  english_tokens %>%
  filter(speaker=="child") %>% #remove data from parents
  group_by(age) %>% #group by age and compute the total number of children in each age bin
  mutate(total_children=length(unique(target_child_id))) %>%
  group_by(negation_type, age, target_child_id, total_children) %>% #for each child, count their negation in each bin
  summarize(count = n()) %>%
  filter(count>2) %>% #remove children that had fewer than 5 tokens in the age bin
  group_by(negation_type,age) %>%
  mutate(producing_children=length(unique(target_child_id)), pc = producing_children/total_children)

proportion_children %>%
  filter(negation_type!="pos")%>%
  ggplot(aes(age, pc)) +
  geom_bar(stat="identity") +
  facet_wrap(.~negation_type)
```

```{r negationRelativeFrequency, fig.env="figure", fig.width=7, fig.height=2.5, fig.cap="Relative frequency of the response particle *no*, verb phrase negation *not*, and its contracted form *n't*"}
negation_relfreq <- readRDS("processed_data/category_table") 

nonotnt <- negation_relfreq %>%
  filter(category!="other", category!="quantifier", category!="adverb", category!="derivational", polarity=="negative")
  
nonotnt$category <- fct_recode(nonotnt$category, `aux-n't` = "auxiliary")
nonotnt$category <- fct_relevel(nonotnt$category, "no", "not")

  ggplot(nonotnt, aes(age,ppt, shape = speaker, color=speaker)) +
  geom_point(stat="identity", aes(shape = speaker, color=speaker)) +
  geom_line() +
  facet_wrap(.~category, scale="free_y") +
  scale_x_continuous(breaks=seq(0, 72, by = 6)) +
  theme_classic() +
  theme(text = element_text(size=10, family="Times"))
```


# Adult Production Permutation Model

```{r}
hist(filter(eng_utterances, speaker=="child", num_tokens<15)$num_tokens, breaks=15)
hist(filter(eng_utterances, speaker=="parent", num_tokens<15)$num_tokens, breaks=15)

nrow(filter(eng_utterances, speaker=="child"))
nrow(filter(eng_utterances, speaker=="parent"))

bottleneck <-
  eng_utterances %>%
  filter(speaker=="child") %>%
  mutate(child_num_tokens = num_tokens) %>%
  select(age, target_child_id, child_num_tokens)

permuterance <-
  eng_utterances %>%
  filter(speaker=="parent") %>%
  left_join(bottleneck, by=c("age", "target_child_id")

permuterance_relfreqs <-
  permuterance %>%
  group_by(speaker, age) %>%
  summarize(no= sum(no)/sum(num_tokens)*1000, not=sum(not)/sum(num_tokens)*1000, nt=sum(nt)/sum(num_tokens)*1000) %>%
  gather("category", "ppt", no:nt)

permuterance_relfreqs %>%
  filter(ppt!=0) %>%
  ggplot(aes(age, ppt)) +
  geom_point(aes(color=category, shape=category)) + 
  geom_line(aes(color=category)) +
  facet_grid(.~speaker) +
  theme_classic()
```



