---
title: 'Negation Production Processing Script: Study 2'
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Documents/LDS/Production_Study/negation_production")
library(tidyverse)
library(magrittr)
library(feather)
library(binom)
library(childesr)
library(lubridate)
library(tm)
library(knitr)
library(kableExtra)
```


```{r, include= FALSE}

# english_negation <- get_utterances(language = "eng", age = c(0,24), role = "Target_Child")
# all_utterances <- english_negation
# 
# 
# count_types <- function(string){
#  n <- nlevels(as.factor(unlist(strsplit(string, " "))))
#  return(n)
# }
# 
# count_tokens <- function(string){
#  n <- unlist(strsplit(string, " "))
#  return(length(n))
# }
# 
# english_negation <-
#  english_negation %>%
#  filter(grepl('^no | no | no$|^not | not | not$|n\'t', gloss)) %>%
#  filter(!grepl('not yet transcribed need to fix', gloss)) %>%
#  mutate(clean_utterances = removeWords(gloss, c("xxx ?", "yyy ?", "um ?", "uh ?", "ah ?"))) %>%
#  mutate(can = grepl('^can | can | can$|cannot|can\'t', clean_utterances)) %>%
#  mutate(do = grepl('^do | do | do$|don\'t|does|doesn\'t', clean_utterances)) %>%
#  mutate(want = grepl('^want | want | want$', clean_utterances))%>%
#  mutate(n_types_clean = unlist(lapply(clean_utterances, count_types))) %>%
#  mutate(n_tokens_clean = unlist(lapply(clean_utterances, count_tokens))) %>%
#  mutate(type_token_ratio = n_types_clean/n_tokens_clean)
# 
# 
# write.csv(english_negation, file="raw_data/eng_negation_undertwo.csv")

```

```{r cleanup, include=FALSE}
undertwo_neg <- read.csv("raw_data/eng_negation_undertwo.csv")

# Modifications to part_of_speech column
condense_pos <- function (gloss, pos) 
{
  # Replace POS with 'no' where relevant
  if (pos == "") {return(pos)}
  no_list = list()
  gloss_list = strsplit(gloss," ")
  for (g in 1:length(gloss_list[[1]])) {
    if (gloss_list[[1]][g] == "no"){
      no_list <- append(unlist(no_list), g)
    }
  }
  pos_list <- unlist(strsplit(pos, " "))
  pos_list[unlist(no_list)] <- "no"
  result <- removeWords(paste(pos_list, collapse = " "), "NA")
  
  # Condense POS to major categories
  result <- result %>%
    str_replace_all('pro:\\w*|n:prop', 'n') %>% 
    str_replace_all('neg', 'not') %>% 
    str_replace_all('adv:\\w*', 'adv') %>% 
    str_replace_all('det:\\w*', 'det')
  
  
  # Condense repeated 'no'
  result <- gsub(' (no )+', ' no ', result) 
  result <- gsub(' (no ?)+$', ' no', result)
  result <- gsub('^(no )+', 'no ', result)
  result <- gsub('^(no ?)+$', 'no', result)
  
  return (result)
}


undertwo_neg$clean_utterances <- as.character(undertwo_neg$clean_utterances)
undertwo_neg$part_of_speech <- as.character(undertwo_neg$part_of_speech)

# Add condensed part of speech column to reflect simplifications
undertwo_neg <- undertwo_neg %>%
  mutate(condensed_p_o_s = 
           mapply(condense_pos, clean_utterances, part_of_speech))

# Fix miscodes
replacements <- 
  list(
    list("no v", "no tickle", "no bite", "no push", "no works", "no hold", "no play", "no touch", "no look", "no help", "no see"),
    list("no v n", "no no like Winnie+Pooh", "no like sweet+corn", "no want din+dins", "no go bathroom", "no touch Lastname",
         "no pop weasel", "no hold it", "no hug baby", "no brush hair", "no pinch that","no ride horsie", "no stop it",
         "no help me", "no spin it", "no play toys"), 
    list("co no co", "quack_quack no peep_peep_peep"),
    list("n no", "horsie no"), 
    list("n mod v prep", "hh I can’t sing with"), 
    list("n mod v n", "I can’t go living+room"), 
    list("no n v", "no Mommy Humm", "no mummy sneeze", "no everybody fall", "no hair wash"), 
    list("no", "no"), 
    list("not adj", "not asleep"), 
    list("no n n v", "no Mommy baby seep"),
    list("no n v n", "no Abi play toys", "no everybody play toys", "no mummy check nappie", "no Papa tie it"),
    list("no v n n", "no hold daddy hand", "no dump Baura puzzle", "no spin it mom", "no pinch that monkey"), 
    list("mod v n adv", "can’t put him back"), 
    list("no n prep n", "no no piece a paper"), 
    list("not det n n", "not a cat station+cat"),
    list("not adj n", "not green one"), 
    list("not n", "not baby", "not lunch", "not diaper", "not blanket"), 
    list("n no v n", "man no taste it")
)
  
fix <- function(df, replacements) 
{
  for (i in 1:length(replacements))
  {
    for (j in 2:length(replacements[[i]]))
    {
      df$condensed_p_o_s[df$clean_utterances == replacements[[i]][[j]]] <- replacements[[i]][[1]]
    }
  }
  return(df)
}

undertwo_neg <- fix(undertwo_neg, replacements)

constructions <- plyr::count(undertwo_neg$condensed_p_o_s) %>%
  arrange(desc(freq))

filtered_constructions <- constructions %>%
  filter(freq>10, x!="no", x!="")

```

Total # of utterances: `r nrow(undertwo_neg)`

|Abbreviation     | Meaning       |
|-----------------|---------------|
| adj             | adjective     |
| adv             | adverb        |
| co              | ???           |
| det             | determiner    |
| mod             | modal         | 
| n               | noun (includes pronouns and proper nouns) |
| part            | participle    |
| prep            | preposition   |
| qn              | quantifier    |
| v               | verb          |


| Construction    | Frequency | Examples                           | 
|-----------------|:-----------:|------------------------------------|
| no n            | `r filter(undertwo_neg, condensed_p_o_s == "no n")$clean_utterances %>% length()` | no baby, no horsie, no juice  | 
| co no           | `r filter(undertwo_neg, condensed_p_o_s == "co no")$clean_utterances %>% length()`| oh no, okay no, yeah no |
| mod v n         | `r filter(undertwo_neg, condensed_p_o_s == "mod v n")$clean_utterances %>% length()` | don't touch that, don't fit car, can't open it|           
| n mod v n       | `r filter(undertwo_neg, condensed_p_o_s == "n mod v n")$clean_utterances %>% length()` | I don't want juice, I can't find duck, cow won't go Mama| 
| mod             | `r filter(undertwo_neg, condensed_p_o_s == "mod")$clean_utterances %>% length()` | don't, can't                       |                      
| n mod v         | `r filter(undertwo_neg, condensed_p_o_s == "n mod v")$clean_utterances %>% length()` | I don't know, they don't fit, I can't open|               
| mod v           | `r filter(undertwo_neg, condensed_p_o_s == "mod v")$clean_utterances %>% length()` | don't go, doesn't work, can't catch|                      
| n mod           | `r filter(undertwo_neg, condensed_p_o_s == "n mod")$clean_utterances %>% length()` | I can't, Jane doesn't, mine didn't |                      
| no qn n         | `r filter(undertwo_neg, condensed_p_o_s == "no qn n")$clean_utterances %>% length()` | no more stuff, no other pennies, no like doggy  |    
| no v            | `r filter(undertwo_neg, condensed_p_o_s == "no v")$clean_utterances %>% length()` | no read, no pooped, no want    |                      
| no n n          | `r filter(undertwo_neg, condensed_p_o_s == "no n n")$clean_utterances %>% length()` | no baby bubble, no this one, no you pencil|              
| no v n          | `r filter(undertwo_neg, condensed_p_o_s == "no v n")$clean_utterances %>% length()` | no sit chair, no have cracker, no eat milk |              
| not n           | `r filter(undertwo_neg, condensed_p_o_s == "not n")$clean_utterances %>% length()` | not noodle, not home, not Toffer's    |                      
| no n v n        | `r filter(undertwo_neg, condensed_p_o_s == "no n v n")$clean_utterances %>% length()` | no car go car, no comb want brush, no mummy do it |       
| not adv         | `r filter(undertwo_neg, condensed_p_o_s == "not adv")$clean_utterances %>% length()` | not today, not there, not yet  |                      
| no adv          | `r filter(undertwo_neg, condensed_p_o_s == "no adv")$clean_utterances %>% length()` | no there, no up, no behind         |                      
| n no            | `r filter(undertwo_neg, condensed_p_o_s == "n no")$clean_utterances %>% length()` | mammy no, apple no, train no       |                      
| no co           | `r filter(undertwo_neg, condensed_p_o_s == "no co")$clean_utterances %>% length()` | no no, no yes, no thanks |
| mod v n n       | `r filter(undertwo_neg, condensed_p_o_s == "mod v n n")$clean_utterances %>% length()` | don't eat skin part, don't touch that book, can't find bunny rabit  | 
| no n v          | `r filter(undertwo_neg, condensed_p_o_s == "no n v")$clean_utterances %>% length()` | no mail came, no mummy go, no I wanted|    
| no prep n       | `r filter(undertwo_neg, condensed_p_o_s == "no prep n")$clean_utterances %>% length()` | no in tree, no with this, no like bubbles|               
| not adj         | `r filter(undertwo_neg, condensed_p_o_s == "not adj")$clean_utterances %>% length()` | not good, not dirty, not real      |                      
| n mod v prep    | `r filter(undertwo_neg, condensed_p_o_s == "n mod v prep")$clean_utterances %>% length()` | I don't want to, Laura don't want to|                     
| n v no          | `r filter(undertwo_neg, condensed_p_o_s == "n v no")$clean_utterances %>% length()` | doll sit no, I take no, Joana say no|                     
| mod n           | `r filter(undertwo_neg, condensed_p_o_s == "mod n")$clean_utterances %>% length()` | can't tea, don't Mommy, didn't I   |                      
| n no v n        | `r filter(undertwo_neg, condensed_p_o_s == "n no v n")$clean_utterances %>% length()` | man no taste it, I no do it   |                      
| not det n       | `r filter(undertwo_neg, condensed_p_o_s == "not det n")$clean_utterances %>% length()` | not my car, not a cat, not a white |                      
| not v n         | `r filter(undertwo_neg, condensed_p_o_s == "not v n")$clean_utterances %>% length()` | not eat rocks, not have coffee, not green one|            
| no det n        | `r filter(undertwo_neg, condensed_p_o_s == "no det n")$clean_utterances %>% length()` | no my turn, no a cow, no the hammer|                    
| mod v adv       | `r filter(undertwo_neg, condensed_p_o_s == "mod v adv")$clean_utterances %>% length()` | can't come in, don't fall down, doesn't fit here|         
| no n adv        | `r filter(undertwo_neg, condensed_p_o_s == "no n adv")$clean_utterances %>% length()` | no jacket on, no monkey there, no lie down  |            
| n no n          | `r filter(undertwo_neg, condensed_p_o_s == "n no n")$clean_utterances %>% length()` | straws no straws, Mommie no mail, there's no towel  |  
| n not n         | `r filter(undertwo_neg, condensed_p_o_s == "n not n")$clean_utterances %>% length()` | pizza not pizza, pizza not apples, that not knees|        
| no adj          | `r filter(undertwo_neg, condensed_p_o_s == "no adj")$clean_utterances %>% length()` | no happy, no pretend, no brown     |                      
| no part         | `r filter(undertwo_neg, condensed_p_o_s == "no part")$clean_utterances %>% length()` | no bouncing, no eating, no hurting |  



| Category           | Frequency  | Examples                                              |
|-----------------------------|:----------:|-------------------------------------------------------|
| Do                 |  `r length(filter(undertwo_neg, do)$clean_utterances)`   | don't like it, I don't feel good, no do it Mummy|
| Ability (can)      |  `r length(filter(undertwo_neg, can)$clean_utterances)`  | I can't find a blankie, can't eat, no no I can not do |
| Desire (want)      |  `r length(filter(undertwo_neg, want)$clean_utterances)` | Don't want to, no comb want brush, no want the baby|
| Don't want         |  `r filter(undertwo_neg, grepl('don\'t want', clean_utterances))$clean_utterances %>% length()`| I don't want juice, don't want it, Laura don't want to|
| Don't know         |   `r filter(undertwo_neg, grepl('don\'t know', clean_utterances))$clean_utterances %>% length()` | I don't know, I don't know name of them, you don't know how zip zip up |
| Don't like         |   `r filter(undertwo_neg, grepl('don\'t like', clean_utterances))$clean_utterances %>% length()` | I don't like it, don't like toast, I don't like|
| No [simple thing]  |  `r filter(undertwo_neg, grepl('^no [a-z]+$', condensed_p_o_s) & condensed_p_o_s != "no no")$clean_utterances %>% length()`| no go, no stupid, no now, no baby|
| No more            |  `r filter(undertwo_neg, grepl('( |^)no more( |$)', clean_utterances))$clean_utterances %>% length()` | no more, no more corn, there's no more time |
| Not [simple thing] |  `r filter(undertwo_neg, grepl('^not [a-z]+$', condensed_p_o_s))$clean_utterances %>% length()`| not bad, not write, not quite|
| Presentential (Optimistic) | `r filter(undertwo_neg, grepl('^no n v ?.*', condensed_p_o_s))$clean_utterances %>% length()` | no daddy eat, no eat lunch, no in tree, no Toto here|
| Presentential (Pessimistic)| `r filter(undertwo_neg, condensed_p_o_s %in% c("no n v", "no n v n"))$clean_utterances %>% length()` | no Mommy read, no I wanted, no baby bite me, no water go|

``` {r, echo=FALSE}

table1 <- data.frame("Category" = c("Can", "Want", "Do","Don't", "Don't want", "Don't know", "Don't like", "No _", "No more", "Not _", "Presentential (Pessimistic)", "Presentential (Optimistic)"), 
                     "Frequency" = c(length(filter(undertwo_neg, can)$clean_utterances), 
                                     length(filter(undertwo_neg, want)$clean_utterances),
                                     length(filter(undertwo_neg, do)$clean_utterances), 
                                     filter(undertwo_neg, grepl('don\'t', clean_utterances))$clean_utterances %>% length(),
                                     filter(undertwo_neg, grepl('don\'t want', clean_utterances))$clean_utterances %>% length(),
                                     filter(undertwo_neg, grepl('don\'t know', clean_utterances))$clean_utterances %>% length(),
                                     filter(undertwo_neg, grepl('don\'t like', clean_utterances))$clean_utterances %>% length(),
                                     filter(undertwo_neg, grepl('^no [a-z]+$', condensed_p_o_s) & condensed_p_o_s != "no no")$clean_utterances %>% length(),
                                     filter(undertwo_neg, grepl('( |^)no more( |$)', clean_utterances))$clean_utterances %>% length(),
                                     filter(undertwo_neg, grepl('^not [a-z]+$', condensed_p_o_s))$clean_utterances %>% length(),
                                     filter(undertwo_neg, grepl('^no n v ?.*', condensed_p_o_s))$clean_utterances %>% length(),
                                     filter(undertwo_neg, grepl('^no v ?.*', condensed_p_o_s))$clean_utterances %>% length()), 
                     "Examples" = c("I can't find a blankie, can't eat, no no I can not do", 
                                    "Don't want to, no comb want brush, no want the baby",
                                    "don't like it, I don't feel good, no do it Mummy", 
                                    "don't like it, no don't, I don't know",
                                    "I don't want juice, don't want it, Laura don't want to",
                                    "I don't know, I don't know name of them, you don't know how zip zip up",
                                    "I don't like it, don't like toast, I don't like",
                                    "no go, no stupid, no now, no baby",
                                    "no more, no more corn, there's no more time",
                                    "not bad, not write, not quite",
                                    "no Mommy read, no baby bite me, no you wait there for me", 
                                    "no have black, no give me that, no bite"))

kable(table1, "latex") %>%
  kable_styling(font_size = 12) %>%
  #pack_rows("Don't", 4, 6); makes a header for rows 4-6
  add_indent(c(4, 5, 6, 7)) %>%
  row_spec(3, bold = T) %>%
  column_spec(3, width = "20em")

```


``` {r, include=FALSE}
#filters for 'no [anything], excluding repeated 'no'; 1183 utterances (33.8%)
filter(undertwo_neg, grepl('^no .*', condensed_p_o_s) & !grepl('^(no ?)+$', condensed_p_o_s))$clean_utterances %>% length()

#filters for 'no n [anything]', excluding repeated 'no'; 236 utterances (6.75%)
filter(undertwo_neg, grepl('^no n.*', condensed_p_o_s) & !grepl('^(no ?)+$', condensed_p_o_s))$clean_utterances %>% length()

# filters for 'do not'; 22 utterances (!!!!) vs. 602 'don't' utterances
filter(undertwo_neg, grepl('( |^)do not( |$)', clean_utterances))$clean_utterances
# Same with 'can not'; 2 utterances vs 208 'can't' utterances


```


```{r analysis, include=FALSE}

# Utterance count, binned by month

breaks <- 0:24
tags <- c("0-1", "1-2", "2-3", "3-4", "4-5", "5-6", "6-7", "7-8", "8-9", "9-10", "10-11", "11-12", "12-13", "13-14", "14-15", "15-16", "16-17", "17-18", "18-19", "19-20", "20-21", "21-22", "22-23", "23-24")

binned <- cut(undertwo_neg$target_child_age, 
              breaks = breaks,
              include.lowest = TRUE,
              labels = tags)

ggplot(as_tibble(binned), aes(x = value)) +
  geom_bar() +
  labs(x = "Age Bin", y = "Utterance Count")

# Participant information

# 173
undertwo_neg$speaker_id %>% unique() %>% length()

speaker_counts <- undertwo_neg %>% count(speaker_id)

# Proportion of total (0, 24) utterances that include negation by age bin (maybe split bar graph?)
#all_utterances
# Proportion of children in each age bin producing negation
# Proportion of utterances that include negation by child (?)
```




