---
title: "Negation Production Processing Script: Study 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

english_negation <- get_utterances(language = "eng", age = c(0,24), role = "Target_Child")


count_types <- function(string){
  n <- nlevels(as.factor(unlist(strsplit(string, " "))))
  return(n)
}

count_tokens <- function(string){
  n <- unlist(strsplit(string, " "))
  return(length(n))
}

english_negation <-
  english_negation %>%
  filter(grepl('^no | no | no$|^not | not | not$|n\'t', gloss)) %>%
  filter(!grepl('not yet transcribed need to fix', gloss)) %>%
  mutate(clean_utterances = removeWords(gloss, c("xxx ?", "yyy ?", "um ?", "uh ?", "oh ?", "ah ?", "n ?"))) %>%
  mutate(can = grepl('^can | can | can$|cannot|can\'t', clean_utterances)) %>%
  mutate(do = grepl('^do | do | do$|don\'t', clean_utterances)) %>%
  mutate(want = grepl('^want | want | want$', clean_utterances))%>%
  mutate(n_types_clean = unlist(lapply(clean_utterances, count_types))) %>%
  mutate(n_tokens_clean = unlist(lapply(clean_utterances, count_tokens))) %>%
  mutate(type_token_ratio = n_types_clean/n_tokens_clean)

write.csv(english_negation, file="raw_data/eng_negation_undertwo.csv")

```

```{r cleanup}
undertwo_neg <- read.csv("raw_data/eng_negation_undertwo.csv")

undertwo_neg[order(undertwo_neg$target_child_age),]

undertwo_neg <- undertwo_neg %>%
  mutate(condensed_p_o_s = 
           (part_of_speech %>%
              str_replace_all(c('co|qn'), 'no') %>%
              str_replace_all(c('pro:\\w*|n:prop'), 'n') %>%
              str_replace_all('neg', 'not') %>%
              str_replace_all('adv:\\w*', 'adv') %>%
              str_replace_all('det:\\w*', 'det')
             )
           )
constructions <- plyr::count(undertwo_neg$condensed_p_o_s) %>%
  arrange(desc(freq))

filtered_constructions <- constructions %>%
  filter(freq>10, x!="no", x!="no no", x!="no no no", x!="no no no no", x!="")

```

```{r tables}

length(filter(undertwo_neg, can)$clean_utterances)
length(filter(undertwo_neg, do)$clean_utterances)
length(filter(undertwo_neg, want)$clean_utterances)

filter(undertwo_neg, grepl('don\'t want', clean_utterances))$clean_utterances %>%
  length()

filter(undertwo_neg, condensed_p_o_s == "no n")$clean_utterances %>%
  length()

no_simple <- c("no adv", "no adj", "no v")

filter(undertwo_neg, condensed_p_o_s %in% no_simple)$clean_utterances %>%
  length()

filter(undertwo_neg, condensed_p_o_s == "not n")$clean_utterances %>%
  length()

not_simple <-  c("not adv", "not adj", "not v")

filter(undertwo_neg, condensed_p_o_s %in% not_simple)$clean_utterances %>%
  length()

optimistic <- c("no n v", "no n adv", "no prep n", "no v n")

filter(undertwo_neg, condensed_p_o_s %in% optimistic)$clean_utterances %>%
  length()
```




Total # of utterances: 3495

|Abbreviation     | Meaning       |
|-----------------|---------------|
| n               | noun (includes pronouns and proper nouns) |
| v               | verb          |
| adj             | adjective     |
| adv             | adverb        |
| prep            | preposition   |
| det             | determiner    |
| part            | participle    |
| mod             | modal         | 


| Construction    | Frequency | Examples                           | Negation Categories  |
|-----------------|-----------|------------------------------------|----------------------|
| no n            | 455       | no baby, no horsie, no juice       | Existence, Desire, Metalinguistic? |
| mod v n         | 142       | don't touch that, don't fit car, can't open it|           |
| n mod v n       | 90        | I don't want juice, I can't find duck, cow won't go Mama| |
| no no n  OOPS   | 83        | PROBLEM HERE no more stuff, *no car, no no pram, #no like doggy  |    |
| mod             | 74        | don't, can't                       |                      |
| n mod v         | 73        | I don't know, they don't fit, I can't open|               |
| mod v           | 67        | don't go, doesn't work, can't catch|                      |
| no n n          | 63        | no baby bubble, no this one, *no brush hair|              |
| n mod           | 55        | I can't, Jane doesn't, mine didn't |                      |
| not n           | 39        | not noodle, not home, not right    |                      |
| no v            | 38        | no read, no pooped,  *no zipper    |                      |
| no v n          | 33        | no sit chair, no have cracker, no eat milk |              |
| not adv         | 30        | not today, not there, *not asleep  |                      |
| n no            | 25        | mammy no, apple no, train no       |                      |
| no adv          | 24        | no there, no up, no behind         |                      |
| no n n n        | 24        | no night night moon, no mummy this side, *no hold daddy hand |                |
| mod v n n       | 22        | don't eat skin part, (?) don't touch that book, *can't find bunny rabit  | |
| no prep n       | 19        | no in tree, no with this, *no like bubbles|               |
| n mod v prep    | 17        | I don't want to, Laura don't want to|                     |
| mod n           | 15        | can't tea, don't Mommy, didn't I   |                      |
| n v no          | 15        | doll sit no, I take no, Joana say no|                     |
| no n v n        | 15        | no car go car, no comb want brush, no mummy do it |       |
| not adj         | 15        | not good, not dirty, not real      |                      |
| not det n       | 15        | not my car, not a cat, not a white |                      |
| not v n         | 15        | not eat rocks, not have coffee, not green one|            |
| no det n        | 13        | no my turn, no a cow, no the hammer|                      |
| not v           | 13        | not shush, not work, not blanket   |                      |
| mod v adv       | 12        | can't come in, don't fall down, doesn't fit here|         |
| no n adv        | 12        | no jacket on, no monkey there, *no lie down  |            |
| no n v          | 12        | no mail came, no mummy go, no I wanted |                  |
| no part         | 12        | no bouncing, no eating, no hurting |                      |
| v no n n        | 12        | man no taste it (???)              |                      |
| n not n         | 11        | pizza not pizza, pizza not apples, that not knees|        |
| no adj          | 11        | no happy, no pretend, no brown     |                      |



| Semantic Category  | Frequency  | Examples  |
|--------------------|------------|-----------|
| Do                 |  666       | don't like it, I don't feel good, no do it Mummy|
| No noun            |  455       | no baby, no horsie, no juice  |
| Ability (can)      |  212       | I can't find a blankie, can't eat, no no I can not do |
| Desire (want)      |  148       | Don't want to, no comb want brush, no want the baby|
| Don't want         |  106       | I don't want juice, don't want it, Laura don't want to|
| No [simple thing]  |  73        | no go, no stupid, no now|
| Not [simple thing] |  58        | not bad, not write, not quite|
| Not noun           |  39        | not noodle, not Danny, not right|
| Presentential (Optimistic) | 76 | |
| Presentential (Pessimistic)| 12 | no Mommy read, no I wanted, no I sweep, no water go|

Notes on Semantic Categories:
- [simple thing] : adj, adv, v
- Presentential (Optimistic) = no n v, no n adv, no prep n, no v n
- Presentential (Pessimistic) = no n v
  - all of these could be read as pronominal


