---
title: 'Negation Production Processing Script: Study 2'
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Documents/LDS/Production_Study/negation_production")
library(tidyverse)
library(magrittr)
library(feather)
library(binom)
library(childesr)
library(lubridate)
library(tm)
```


```{r, include= FALSE}

# english_negation <- get_utterances(language = "eng", age = c(0,24), role = "Target_Child")
# 
# 
# count_types <- function(string){
#  n <- nlevels(as.factor(unlist(strsplit(string, " "))))
#  return(n)
# }
# 
# count_tokens <- function(string){
#  n <- unlist(strsplit(string, " "))
#  return(length(n))
# }
# 
# english_negation <-
#  english_negation %>%
#  filter(grepl('^no | no | no$|^not | not | not$|n\'t', gloss)) %>%
#  filter(!grepl('not yet transcribed need to fix', gloss)) %>%
#  mutate(clean_utterances = removeWords(gloss, c("xxx ?", "yyy ?", "um ?", "uh ?", "ah ?"))) %>%
#  mutate(can = grepl('^can | can | can$|cannot|can\'t', clean_utterances)) %>%
#  mutate(do = grepl('^do | do | do$|don\'t|does|doesn\'t', clean_utterances)) %>%
#  mutate(want = grepl('^want | want | want$', clean_utterances))%>%
#  mutate(n_types_clean = unlist(lapply(clean_utterances, count_types))) %>%
#  mutate(n_tokens_clean = unlist(lapply(clean_utterances, count_tokens))) %>%
#  mutate(type_token_ratio = n_types_clean/n_tokens_clean)
# 
# 
# write.csv(english_negation, file="raw_data/eng_negation_undertwo.csv")

```

```{r cleanup, include=FALSE}
undertwo_neg <- read.csv("raw_data/eng_negation_undertwo.csv")

# uh no idea what this does??
#undertwo_neg[order(undertwo_neg$target_child_age),]

# Takes in a gloss string and a part of speech string and returns a part of speech string with "no" in places
# corresponding to "no" in the gloss
set_nos <- function(gloss, pos) {
  if (pos == "") {return(pos)}
  no_list = list()
  gloss_list = strsplit(gloss," ")
  for (g in 1:length(gloss_list[[1]])) {
    if (gloss_list[[1]][g] == "no"){
      no_list <- append(unlist(no_list), g)
    }
  }
  pos_list <- unlist(strsplit(pos, " "))
  pos_list[unlist(no_list)] <- "no"
  return(paste(pos_list, collapse = " "))
}

# Modifications to part_of_speech column
replace_f <- function (gloss, pos) 
{
  result <- pos %>%
    str_replace_all('pro:\\w*|n:prop', 'n') %>% 
    str_replace_all('neg', 'not') %>% 
    str_replace_all('adv:\\w*', 'adv') %>% 
    str_replace_all('det:\\w*', 'det')
  
  return (set_nos(gloss, result))
}


undertwo_neg$clean_utterances <- as.character(undertwo_neg$clean_utterances)
undertwo_neg$part_of_speech <- as.character(undertwo_neg$part_of_speech)

# Add condensed part of speech column to reflect simplifications
undertwo_neg <- undertwo_neg %>%
  mutate(condensed_p_o_s = 
           mapply(replace_f, clean_utterances, part_of_speech))

constructions <- plyr::count(undertwo_neg$condensed_p_o_s) %>%
  arrange(desc(freq))

filtered_constructions <- constructions %>%
  filter(freq>10, x!="no", x!="no no", x!="no no no", x!="no no no no", x != "no no no no no", x!="")

```

Total # of utterances: `r nrow(undertwo_neg)`

|Abbreviation     | Meaning       |
|-----------------|---------------|
| adj             | adjective     |
| adv             | adverb        |
| co              | ???           |
| det             | determiner    |
| mod             | modal         | 
| n               | noun (includes pronouns and proper nouns) |
| part            | participle    |
| prep            | preposition   |
| qn              | quantifier    |
| v               | verb          |


| Construction    | Frequency | Examples                           | 
|-----------------|-----------|------------------------------------|
| no n            | `r filter(undertwo_neg, condensed_p_o_s == "no n")$clean_utterances %>% length()` | no baby, no horsie, no juice  | 
| co no           | `r filter(undertwo_neg, condensed_p_o_s == "co no")$clean_utterances %>% length()`| oh no, okay no, yeah no |
| mod v n         | `r filter(undertwo_neg, condensed_p_o_s == "mod v n")$clean_utterances %>% length()` | don't touch that, don't fit car, can't open it|           
| n mod v n       | `r filter(undertwo_neg, condensed_p_o_s == "n mod v n")$clean_utterances %>% length()` | I don't want juice, I can't find duck, cow won't go Mama| 
| mod             | `r filter(undertwo_neg, condensed_p_o_s == "mod")$clean_utterances %>% length()` | don't, can't                       |                      
| n mod v         | `r filter(undertwo_neg, condensed_p_o_s == "n mod v")$clean_utterances %>% length()` | I don't know, they don't fit, I can't open|               
| mod v           | `r filter(undertwo_neg, condensed_p_o_s == "mod v")$clean_utterances %>% length()` | don't go, doesn't work, can't catch|                      
| no n n          | `r filter(undertwo_neg, condensed_p_o_s == "no n n")$clean_utterances %>% length()` | no baby bubble, no this one, *no brush hair|              
| n mod           | `r filter(undertwo_neg, condensed_p_o_s == "n mod")$clean_utterances %>% length()` | I can't, Jane doesn't, mine didn't |                      
| no qn n         | `r filter(undertwo_neg, condensed_p_o_s == "no qn n")$clean_utterances %>% length()` | no more stuff, *no car, no no pram, no like doggy  |    
| no co           | `r filter(undertwo_neg, condensed_p_o_s == "no co")$clean_utterances %>% length()` | no no, no yes, no look (??), no thanks |
| not n           | `r filter(undertwo_neg, condensed_p_o_s == "not n")$clean_utterances %>% length()` | not noodle, not home, not right    |                      
| no v            | `r filter(undertwo_neg, condensed_p_o_s == "no v")$clean_utterances %>% length()` | no read, no pooped,  *no zipper    |                      
| no v n          | `r filter(undertwo_neg, condensed_p_o_s == "no v n")$clean_utterances %>% length()` | no sit chair, no have cracker, no eat milk |              
| not adv         | `r filter(undertwo_neg, condensed_p_o_s == "not adv")$clean_utterances %>% length()` | not today, not there, *not asleep  |                      
| no adv          | `r filter(undertwo_neg, condensed_p_o_s == "no adv")$clean_utterances %>% length()` | no there, no up, no behind         |                      
| no n n n        | `r filter(undertwo_neg, condensed_p_o_s == "no n n n")$clean_utterances %>% length()` | no night night moon, no mummy this side, *no hold daddy hand  |
| mod v n n       | `r filter(undertwo_neg, condensed_p_o_s == "mod v n n")$clean_utterances %>% length()` | don't eat skin part, (?) don't touch that book, *can't find bunny rabit  | 
| n no            | `r filter(undertwo_neg, condensed_p_o_s == "n no")$clean_utterances %>% length()` | mammy no, apple no, train no       |                      
| no no n         | `r filter(undertwo_neg, condensed_p_o_s == "no no n")$clean_utterances %>% length()` | no no mine, no no more, no no mouse |
| no prep n       | `r filter(undertwo_neg, condensed_p_o_s == "no prep n")$clean_utterances %>% length()` | no in tree, no with this, *no like bubbles|               
| n mod v prep    | `r filter(undertwo_neg, condensed_p_o_s == "n mod v prep")$clean_utterances %>% length()` | I don't want to, Laura don't want to|                     
| n v no          | `r filter(undertwo_neg, condensed_p_o_s == "n v no")$clean_utterances %>% length()` | doll sit no, I take no, Joana say no|                     
| mod n           | `r filter(undertwo_neg, condensed_p_o_s == "mod n")$clean_utterances %>% length()` | can't tea, don't Mommy, didn't I   |                      
| no n v n        | `r filter(undertwo_neg, condensed_p_o_s == "no n v n")$clean_utterances %>% length()` | no car go car, no comb want brush, no mummy do it |       
| not adj         | `r filter(undertwo_neg, condensed_p_o_s == "not adj")$clean_utterances %>% length()` | not good, not dirty, not real      |                      
| not det n       | `r filter(undertwo_neg, condensed_p_o_s == "not det n")$clean_utterances %>% length()` | not my car, not a cat, not a white |                      
| not v n         | `r filter(undertwo_neg, condensed_p_o_s == "not v n")$clean_utterances %>% length()` | not eat rocks, not have coffee, not green one|            
| no det n        | `r filter(undertwo_neg, condensed_p_o_s == "no det n")$clean_utterances %>% length()` | no my turn, no a cow, no the hammer|                    
| not v           | `r filter(undertwo_neg, condensed_p_o_s == "not v")$clean_utterances %>% length()` | not shush, not work, not blanket   |                      
| mod v adv       | `r filter(undertwo_neg, condensed_p_o_s == "mod v adv")$clean_utterances %>% length()` | can't come in, don't fall down, doesn't fit here|         
| no n adv        | `r filter(undertwo_neg, condensed_p_o_s == "no n adv")$clean_utterances %>% length()` | no jacket on, no monkey there, *no lie down  |            
| no n v          | `r filter(undertwo_neg, condensed_p_o_s == "no n v")$clean_utterances %>% length()` | no mail came, no mummy go, no I wanted|    
| v no n n        | `r filter(undertwo_neg, condensed_p_o_s == "v no n n")$clean_utterances %>% length()` | man no taste it (???)              |  
| n not n         | `r filter(undertwo_neg, condensed_p_o_s == "n not n")$clean_utterances %>% length()` | pizza not pizza, pizza not apples, that not knees|        
| no adj          | `r filter(undertwo_neg, condensed_p_o_s == "no adj")$clean_utterances %>% length()` | no happy, no pretend, no brown     |                      
| no part         | `r filter(undertwo_neg, condensed_p_o_s == "no part")$clean_utterances %>% length()` | no bouncing, no eating, no hurting |  



| Semantic Category           | Frequency  | Examples                                              |
|-----------------------------|:----------:|-------------------------------------------------------|
| Do                 |  `r length(filter(undertwo_neg, do)$clean_utterances)`   | don't like it, I don't feel good, no do it Mummy|
| Ability (can)      |  `r length(filter(undertwo_neg, can)$clean_utterances)`  | I can't find a blankie, can't eat, no no I can not do |
| Desire (want)      |  `r length(filter(undertwo_neg, want)$clean_utterances)` | Don't want to, no comb want brush, no want the baby|
| Don't want         |  `r filter(undertwo_neg, grepl('don\'t want', clean_utterances))$clean_utterances %>% length()`| I don't want juice, don't want it, Laura don't want to|
| No [simple thing]  |  `r filter(undertwo_neg, grepl('^no [a-z]+$', condensed_p_o_s) & condensed_p_o_s != "no no")$clean_utterances %>% length()`| no go, no stupid, no now, no baby|
| Not [simple thing] |  `r filter(undertwo_neg, grepl('^not [a-z]+$', condensed_p_o_s))$clean_utterances %>% length()`| not bad, not write, not quite|
| Presentential (Optimistic) | `r filter(undertwo_neg, condensed_p_o_s %in% c("no n v", "no n v n", "no n adv", "no prep n", "no v n"))$clean_utterances %>% length()` | no daddy eat, no eat lunch, no in tree, no Toto here|
| Presentential (Pessimistic)| `r filter(undertwo_neg, condensed_p_o_s %in% c("no n v", "no n v n"))$clean_utterances %>% length()` | no Mommy read, no I wanted, no baby bite me, no water go|




